---
title: "sobol_with_ode"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(deSolve)
library(sensitivity)
```

```{r sen}
source("R/dpopgrowth.R")

dpopgrowth

# come up with initial set of sample parameters
# we will assume that we know the initial forest size of 10,
# initial forest size
time = seq(1, 300)
Cinitial = 10
thresh = 50
K = 250
r = 0.01
g = 2

parms = list(K = 250, r = 0.01, g = 2, thresh = 50)
result = ode(y=Cinitial, 
             times=time,
             func=dpopgrowth, # growth function
             parms=parms,
             thresh = thresh)

result <- as.data.frame(result)

# rename 1 to carbon 
ggplot(result, aes(time, `1`)) +
  geom_point()




```




```{r}

Cinitial = 10
thresh = 50
# want to learn about sensitivity to growth rate (r) and carrying capacity 
# set the number of parameters
np=2000

K = rnorm(mean=250, sd=250*0.01, n=np)
r = rnorm(mean=0.01, sd=0.01*0.01, n=np)
g = rnorm(mean=2, sd=2*0.01, n=np)


X1 = cbind.data.frame(K=K, r=r, g=g)

# repeat to get our second set of samples
K = rnorm(mean=250, sd=250*0.01, n=np)
r = rnorm(mean=0.01, sd=0.01*0.01, n=np)
g = rnorm(mean=2, sd=2*0.01, n=np)

X2 = cbind.data.frame(K=K, r=r, g=g)

# fix any negative values and they are not meaningful
X1 = X1 %>% map_df(pmax, 0.0)
X2 = X2 %>% map_df(pmax, 0.0)

# create our sobel object and get sets of parameters for running the model
# set 300 for number of bootstrap replicates

sens_C = sobolSalt(model = NULL, X1, X2, nboot = 300)

# our parameter sets are
head(sens_C$X)

# lets add names 
colnames(sens_C$X) = c("r","K", "g")

head(sens_C$X)

# gets results for 300 years (evaluating every year)
simtimes = seq(from=1, to=300)

parms = list(r=sens_C$X[1,"r"], K=sens_C$X[1,"K"], g=sens_C$X[1,"g"])

result = ode(y=Cinitial,
             thresh = thresh,
             times=simtimes,
             func=dpopgrowth,
             parms=parms)

# head(result)
# colnames(result)=c("time","C")
# 
# # turn it into a data frame
# result = as.data.frame(result)
# 
# ggplot(result, aes(time, C))+
#   geom_point()


```
# Compute our metric for all the parameter sets

```{r}
# gets results for 200 years (evaluating every year)
thresh = 50
simtimes = seq(from=1, to=300)
parms = list(r=sens_P$X[1,"r"], 
             K=sens_P$X[1,"K"], 
             g=sens_P$X[1,"g"])

# run the ODE solver with these conditions
result = ode(y=Cinitial, 
             times=simtimes,
             func=dforestgrowth, # growth function
             parms=parms,
             thresh = thresh)

result = as.data.frame(result)

# rename the column names
colnames(result)=c("time","C")
```


What if we want to run for all parameters

Lets create two additional functions that will help us

* a function that computes the metrics we want

* a function that runs our ode solver and computes the metrics (I call it a wrapper function as it is really just a workflow/wrapper to call ode solver and then compute metrics)
```{r sen2}
# turn computing our metrics into a function

compute_metrics = function(result, thresh) {
  maxpop = max(result$P)
idx = which(result$P > thresh)[1]
idx = ifelse(is.na(idx), length(result$P), idx)
threshyear = result$time[idx]
return(list(maxpop=maxpop, threshyear=threshyear))}

# try it on our first parameter set, and look at when it gets to 100
compute_metrics(result, 100)

# great but we need to apply the ode and this function for all of our parameters



# define a wrapper function to do everything we need - run solver and compute metrics - and send back results for each parameter

# lets make the threshold 90% of carrying capacity

p_wrapper = function(r,K, Pinitial, simtimes, odefunc, metricfunc) {
    parms = list(r=r, K=K)
    result = ode(y=Pinitial, times=simtimes, func=odefunc, parms=parms) 
    colnames(result)=c("time","P")
  # get metrics
  metrics=metricfunc(as.data.frame(result), thresh=100)
  return(metrics)
}

# test
p_wrapper(r=0.01, K=150, Pinitial=3, simtimes=seq(from=1,to=10),
         odefunc=dpopgrowth, metricfunc=compute_metrics)
```



# Next step

Run the wrapper for all parameters and look at results


```{r userwarpper}
# now use pmap as we did before

allresults = as.data.frame(sens_P$X) %>% pmap(p_wrapper, Pinitial=Pinitial, simtimes=simtimes, odefunc=dpopgrowth, metricfunc=compute_metrics)

# extract out results from pmap into a data frame
allres = allresults %>% map_dfr(`[`,c("maxpop","threshyear"))

# maxpop and threshyear
allres

# can put everything together
tmp_all = cbind.data.frame(allres, sens_P$X)
head(tmp_all)

# why does this vary, because we have varied K aka carrying capacity
ggplot(tmp_all, aes(r, threshyear, col=K))+geom_point()


# create boxplots
tmp = allres %>% pivot_longer(cols=everything(),names_to="metric", values_to="value")
ggplot(tmp, aes(metric, value, col=metric))+geom_boxplot()

```

# Compute the sobol indicies for each metric

```{r sen3}
# sobol can only handle one output at a time  - so we will need to do them separately
sens_P_maxpop = sensitivity::tell(sens_P,allres$maxpop)

# first-order indices (main effect without co-variance)
rownames(sens_P_maxpop$S)=c("r","K")
sens_P_maxpop$S

# total sensitivity index -note that this partitions the output variance 
rownames(sens_P_maxpop$T)=c("r","K")
sens_P_maxpop$T


# create another one for max year; tell 
sens_P_threshyear = sensitivity::tell(sens_P,allres$threshyear)
# first-order indices (main effect without co-variance) - k uncertainty is more important for max
rownames(sens_P_threshyear$S)=c("r","K")
sens_P_threshyear$S

# total sensitivity index -note that this partitions the output variance - so values sum to 1
rownames(sens_P_threshyear$T)=c("r","K")
sens_P_threshyear$T

```

